---
title: "Exercises"
output:
  word_document: default
  pdf_document: default
---


## Exercise 1:

```{r setup, message=FALSE,warning=FALSE}
library(mosaic)
library(tidyverse)
```

Read in "NHANES" dataset and take the adult (age $18+$ subset:
(change the code below so that it points to where you have nhanes.csv saved)
```{r}
nhanes <- read_csv('nhanes.csv')
### only adults
nhanes <- filter(nhanes,age>=18)
```

Consider weights for people with "PreHypertension," or systolic blood pressure between 120 and 140.
```{r}
PreHyp <- filter(nhanes,sbpClass=='Prehypertension')
```

 * Use the formula (as above) to calculate the 90% C.I. for the mean weight for all PreHypertensive people in the population represented by the NHANES survey. 
```{r}
NormalSBP <- filter(nhanes, sbpClass == 'Normal')
# put your answer here
weight <- NormalSBP$weight
sampleMean <- mean(weight)
sampleSD <- sd(weight)
sampleSize <- length(weight)
SE <- sampleSD/sqrt(sampleSize)
alpha <- 1- 0.90 ## proportion of time CI is wrong
critValue <- -qt(alpha/2,df=sampleSize-1)
CI.lower <- sampleMean - critValue*SE #Lower bound of the C.I.
CI.lower
```

 * Use function t.test to calculate the 90% C.I. for the mean weight for all PreHypertensive people in the population represented by the NHANES survey. 
```{r}
t.test(weight,conf.level = 0.9)
# put your answer here
```
* What is the probability that the true population mean is inside the confidence interval you estimated? Does this question even make sense? Why or why not?

The probability is 90%. Yes it makes sense becuase that is the confidence inerval we are taking, it is likely that the mean will be inside.

* If you were to gather a large number of surveys from the same population, using the same methodology of NHANES, and estimate a confidence interval using each survey in the same way, what proportion of those confidence intervals should include the true population mean? 

9/10 should include the true population mean 


# Exercise 2: 

Now study the difference of the mean weight between adults with prehypertension and adults with SPB Stage 1 Hypertension. 

We already have a dataset of prehypertensive people. Here's a dataset of people with Stage 1 Hypertension:
```{r}
stage1 <- filter(nhanes,sbpClass=='Stage 1 Hypertension')
```

Use the formula to calculate the 90% C.I. for the mean weight difference:

* What is the standard error of the difference?
```{r}
w1 <- PreHyp$weight
w2 <- stage1$weight
w1 <- NormalSBP$weight; #Group1: weights for normal SBP;
PreHyp <- filter(nhanes,sbpClass=='Prehypertension')
w2 <- PreHyp$weight; #Group2: weights for prehypertension;

### Calculation by formula: calculate the 95% confidence interval 
#   for the difference of population means
y1 <- mean(w1); #sample mean of group1. 
y2 <- mean(w2); #sample mean of group2.

s1 <- sd(w1); #sample SD
s2 <- sd(w2); 

n1 <- length(w1); #sample size
n2 <- length(w2); #sample size

V1 <- s1^2/n1; #sampling variance of sample mean of group 1.
V2 <- s2^2/n2; #sampling variance of sample mean of group 2.

Vdiff <- V1+V2 # sampling variance of difference of two sample means

SE <- sqrt(Vdiff); #Standard error for the difference of sample means.
SE


```

* What is a 90% CI of the difference?
```{r}
## put your answer here
alpha <- 1- 0.90 ## proportion of time CI is wrong
critValue <- -qt(alpha/2,df=sampleSize-1)
CI.lower <- sampleMean - critValue*SE #Lower bound of the C.I.
CI.lower
```

 * Use function t.test to calculate the 90% C.I. for the mean weight difference.
```{r}
## put your answer here
t.test(w1, w2,conf.level = 0.90); 
```


